Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across \nOntario’s Far North

### 2024-01-18 reports/GnC_report/paper/paper.qmd
fix up the qmd, but I really need to get back into it using the code Eric is using

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '~/Documents/WILDLab/rofcamtrap/')
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
backup_options <- options()
options(scipen = 1, digits = 2)
library(magrittr)
```

@@

RoF_CameraDeploymentLog_WLU <- 
  readxl::read_excel(paste0("analysis/data/raw_data/camera_deployments/",
                            "WLU_Cameras_Log_2023.xlsx")) %>% 
  janitor::clean_names()

cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
  dplyr::filter(!is.na(latitude)) %>% 
  dplyr::filter(!is.na(longitude)) %>% 
  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326", remove=F) %>% 
  dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))

coords <- cams_sf %>% 
  sf::st_coordinates() %>% 
  as.data.frame()

@@

# zones <- 
#   sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
#               quiet=TRUE, ) %>% 
#   sf::st_transform(crs = sf::st_crs(cams_sf))

zones <- 
  sf::st_read("analysis/data/raw_data/ecozones/EcoZone/EcoZone.shp",
              quiet=TRUE, ) %>% 
  sf::st_transform(crs = sf::st_crs(cams_sf))

zones_filter <- zones %>%
  sf::st_filter(cams_sf)

pols <- sf::st_coordinates(zones_filter) %>% 
  as.data.frame()

# regions_labels <- zones_filter %>% 
#   sf::st_drop_geometry() %>% 
# dplyr::bind_cols(data.frame(x = c(-88.5, -84, -88.5),
#                             y = c(55, 52, 52.6)))

regions_labels <- zones_filter %>% 
  sf::st_drop_geometry() %>% 
  dplyr::bind_cols(data.frame(x = c(-84.5, -88.5),
                              y = c(52, 52.6)))

map <- ggmap::get_stamenmap( bbox = c(left = -90, bottom = 50, right = -83, top = 56), 
                             zoom = 9, maptype = "terrain-background")
```

@@

base <- ggmap::ggmap(map) +
  ggplot2::labs(x = "Longitude", y = "Latitude") +
  # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
  ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)

p <- base +
  # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
  #                     size = 2.5, data = regions_labels) +
  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
                      size = 2.5, data = regions_labels) +
  ggplot2::geom_point(data = cams_sf_coords, cex = 1, shape=21, col="black",
                      ggplot2::aes(x=X, y=Y, fill=is_retrieved), 
                      inherit.aes = FALSE) +
  ggplot2::labs(fill = "Camera \nRetrieved")
p
```

@@

# base <- ggmap::ggmap(map) +
#   ggplot2::labs(x = "Longitude", y = "Latitude") +
#   # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
#   ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
# 
# p <- base +
#   # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
#   #                     size = 2.5, data = regions_labels) +
#   ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
#                       size = 2.5, data = regions_labels) +
#   ggplot2::geom_point(data = cams_sf_coords %>% 
#                         dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
#                       col="black", fill = "red",
#                       ggplot2::aes(x=X, y=Y), 
#                       inherit.aes = FALSE) +
#   ggplot2::labs(fill = "Camera \nRetrieved")
# p
```

@@

if (!file.exists("analysis/data/derived_data/annotations.rds") |
    !file.exists("analysis/data/derived_data/annotations_wide.rds")){
  # source("scripts/1_preprocess_annotations.R")
} else {
  # anns <- readRDS("../analysis/data/derived_data/annotations.rds")
  # anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide_treated.rds")
}

if (!file.exists("analysis/data/derived_data/detections.rds")){
  # source("scripts/2_preprocess_detections.R")
} else {
  dets <- readRDS("analysis/data/derived_data/detections.rds")
}

@@

# Hardcoded check
stopifnot(nb_image == 2788375)

exif <- readr::read_csv("analysis/data/raw_data/exif/exif_combined.csv") %>% 
  dplyr::select(-SourceFile) %>% 
  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) %>% 
  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) %>% 
  dplyr::mutate(image_id = stringr::str_sub(image_id, 6)) %>% 
  janitor::clean_names()
```

@@

# Select and distinct down to image-based infos
dets_images <- dets %>%  dplyr::select(image_id, max_confidence, isempty) %>% 
  dplyr::distinct()

# Breakdown of which images triggered a MD detection
dets_tb <- dets_images$isempty %>% table()
dets_tb_sum <- sum(dets_tb)
dets_tb_percent <- dets_tb/dets_tb_sum*100

@@

```{r}
#| label: basics-stats-2
#| echo: false
dets_0 <- dets %>% dplyr::filter(confidence > 0)
dets_0.1 <- dets %>% dplyr::filter(confidence > 0.1)

# Breakdown of detections by categories (animal, person, vehicles)
dets_tb_cat_unfil <- dets_0 %>% dplyr::pull(category_id) %>% table()
dets_tb_cat_unfil_sum <- sum(dets_tb_cat_unfil)
dets_tb_cat_unfil_percent <- dets_tb_cat_unfil/dets_tb_cat_unfil_sum*100

@@

# Breakdown of detections above 0.1 by categories (animal, persons, vehicles)
dets_tb_cat <- dets_0.1 %>% dplyr::pull(category_id) %>% table()
dets_tb_cat_sum <- sum(dets_tb_cat)
dets_tb_cat_percent <- dets_tb_cat/sum(dets_tb_cat)*100

@@

# Histogram of confidence scores above 0.1
dets_0.1 %>% 
  dplyr::mutate(category_id = as.character(category_id)) %>% 
  dplyr::mutate(Type = dplyr::case_match(category_id,
                                         "0" ~ "Empty",
                                         "1" ~ "Animal",
                                         "2" ~ "Person",
                                         "3" ~ "Vehicle")) %>% 
  ggplot2::ggplot() +
  ggplot2::theme_bw() +
  ggplot2::geom_histogram(ggplot2::aes(x=confidence, group=Type, fill=Type),
                          bins = 100) +
  ggplot2::scale_fill_viridis_d() +
  ggplot2::xlab("Confidence Score") +
  ggplot2::ylab("Count") +
  ggplot2::geom_vline(ggplot2::aes(xintercept = 0.1), lty=2, col = "firebrick")
```

@@

# Number of uploaded images
uploaded <- dets %>% dplyr::filter(confidence > 0.1) %>% dplyr::pull(image_id) %>% 
  unique() %>% length()

@@

# Filter for non empty images
anns_wide_noempty <- anns_wide %>%
  dplyr::filter(!is.na(to_name))

saveRDS(anns_wide_noempty, "analysis/data/derived_data/annotations_wide_noempty.rds")

@@

# Deleted predictions
dels <- anns_wide_noempty %>% dplyr::filter(is.na(species))
dels_nb <- nrow(dels)
dels_nb_fmt <- prettyNum(dels_nb, big.mark = ",")
dels_nb_pct <- dels_nb/nrow(anns_wide_noempty)*100

# Should be no manuals here
stopifnot((anns_wide_noempty %>% dplyr::filter(is.na(species)) 
           %>% dplyr::filter(manual)) == 0)

# Actual annotations
true_anns <- anns_wide_noempty %>% dplyr::filter(!is.na(species))
true_anns_nb <- nrow(true_anns)
# true_anns_nb_fmt <- prettyNum(true_anns_nb, big.mark = ",")
# true_anns_nb_pct <- true_anns_nb/processed*100

# Non manuals 
non_mans <- anns_wide_noempty %>% dplyr::filter(!is.na(species)) %>% 
  dplyr::filter(!manual)
non_mans_nb <- nrow(non_mans)
non_mans_nb_fmt <- prettyNum(non_mans_nb, big.mark = ",")
non_mans_nb_pct <- non_mans_nb/nrow(anns_wide_noempty)*100

# Mans
mans <- anns_wide_noempty %>% dplyr::filter(!is.na(species)) %>% 
  dplyr::filter(manual)
mans_nb <- nrow(mans)
mans_nb_fmt <- prettyNum(mans_nb, big.mark = ",")
mans_nb_pct <- mans_nb/nrow(anns_wide_noempty)*100

# adjusted
orig_tb <- anns_wide_noempty$origin %>% table()
orig_tb_pct <- orig_tb/sum(orig_tb)*100
orig_tb_fmt <- prettyNum(orig_tb, big.mark = ",")
```

@@

# Only annotations for animals
anns_wide_animals <- anns_wide_noempty %>% 
  dplyr::filter(species != 'Human') %>% 
  dplyr::filter(species != 'Vehicle') %>% 
  dplyr::mutate(camera_id = id %>% stringr::str_split("_") %>% 
                  purrr::map(~.x[[2]]) %>% unlist())

# Unknowns and bird breakdown
# anns_wide_animals %>% dplyr::filter(species == "Unknown") %>% 
#   dplyr::pull(id) %>% 
#   stringr::str_sub(1,4) %>% table() %>% sort() 
# anns_wide_animals %>% dplyr::filter(stringr::str_detect(species,"Bird")) %>% 
#   dplyr::pull(id) %>% 
#   stringr::str_sub(1,4) %>% table() %>% sort()

@@

anns_wide_animals_joined <- anns_wide_animals %>% 
  dplyr::left_join(exif, by = c("source_file", "image_id"))

anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
  dplyr::filter(species != "Unknown") %>%
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_"))

# anns_wide_animals_joined_dets <- anns_wide_animals %>% 
#   dplyr::right_join(dets_images, by = "image_id")
```

```{r eval=FALSE}
# library(forcats)
library(ggplot2)
# anns_wide_animals_joined_mut %>% 
#   ggplot(aes(x=date_time, y=species)) + geom_point()
# anns_wide_animals_joined_mut %>% 
#   dplyr::filter(species == "Fisher") %>% View()

camera_days <- exif %>% dplyr::group_by(maker_notes_user_label) %>% 
  dplyr::summarise(camera_first_time=min(exif_date_time_original)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_")) %>% 
  dplyr::select(camera_id = maker_notes_user_label, 
                camera_first_day = date_time)

camera_regions <- cams_sf %>% sf::st_join(zones) %>% 
  sf::st_drop_geometry() %>% dplyr::select(camera_id, ECOZONE_NA)

bp <- anns_wide_animals_joined_mut %>%

  dplyr::filter(species != "Fisher") %>% 
  dplyr::left_join(camera_days, by = "camera_id") %>%

  dplyr::select(species, camera_id, 
                date_time, camera_first_day, ymd) %>%

  dplyr::group_by(species, camera_id) %>% 
  dplyr::summarise(first_day = min(date_time),
                   camera_first_day = min(camera_first_day),
                   camera_events = length(unique(ymd))) %>% 
  dplyr::ungroup() %>%

  dplyr::left_join(camera_regions, by = "camera_id") %>%

  dplyr::group_by(species) %>% 
  dplyr::mutate(total_events = sum(camera_events),
                n_cameras = dplyr::n()) %>% 
  dplyr::ungroup() %>%

  # dplyr::filter(n_cameras > 4) %>%
  dplyr::filter(n_cameras > 3) %>%

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) %>%

@@

  scale_x_continuous()+
  labs(x = "Detection time (in days)",
       y = "Species",
       color = "Ecozone") + 
  theme_bw()
ggsave(plot = bp, filename = "analysis/figures/bp.png", width = 10, height = 15)

pts <- anns_wide_animals_joined_mut %>%

  dplyr::filter(species != "Fisher") %>% 
  dplyr::left_join(camera_days, by = "camera_id") %>%

  dplyr::select(species, camera_id, 
                date_time, camera_first_day, ymd) %>%

  dplyr::group_by(species, camera_id) %>% 
  dplyr::summarise(first_day = min(date_time),
                   camera_first_day = min(camera_first_day),
                   camera_events = length(unique(ymd))) %>% 
  dplyr::ungroup() %>%

  dplyr::left_join(camera_regions, by = "camera_id") %>%

  dplyr::group_by(species) %>% 
  dplyr::mutate(total_events = sum(camera_events),
                n_cameras = dplyr::n()) %>% 
  dplyr::ungroup() %>%

  # dplyr::filter(n_cameras > 4) %>%
  dplyr::filter(n_cameras > 3) %>%

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) %>%

@@

  scale_x_continuous()+
  labs(x = "Detection time (in days)",
       y = "Species",
       color = "Ecozone",
       size = "Day counts") + 
  theme_bw()
ggsave(plot = pts, filename = "analysis/figures/pts.png", width = 10, height = 15)
pts
```

```{r eval=FALSE}
for_vic <- anns_wide_animals_joined_mut %>% 
  dplyr::select(id,species,date_time, camera_id) %>% 
  dplyr::left_join(cams_sf %>% dplyr::select(camera_id, plot_id) %>% sf::st_drop_geometry(),
                   by = "camera_id")
tot_camera <- for_vic$camera_id %>% unique() %>% length()
hit_rates <- for_vic %>% dplyr::select(species, camera_id) %>% 
  dplyr::group_by(species) %>% dplyr::summarise(n_cameras=length(unique(camera_id))) %>% 
  dplyr::ungroup() %>% dplyr::mutate(rate = n_cameras/tot_camera*100) %>% 
  dplyr::arrange(dplyr::desc(rate))
# det_history_two_weeks <-

readr::write_csv(for_vic, "../data/derived_data/observations_by_camera_2023Aug15.csv")
readr::write_csv(hit_rates, "../data/derived_data/hit_rates_by_species_2023Aug15.csv")
```

```{r}
#| label: species
#| echo: false
anns_wide_animals_table <- anns_wide_animals %>% 
  dplyr::select(camera_id, species) %>% 
  dplyr::filter(species != "Unknown") %>%
  table() %>% 
  `>`(0) %>% 
  `+`(0) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("camera_id") %>% 
  dplyr::arrange(camera_id)
anns_wide_animals_table_tb <- anns_wide_animals_table %>% 
  dplyr::select(-camera_id)
rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id
```

@@

lu_16 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
lu_17 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")

lu_dat <- readr::read_csv("analysis/data/raw_data/land_use/attr_table_northen_ont_lc.txt") %>% 
  dplyr::mutate(cats = as.factor(code))

cams_sf_buffers_16 <- cams_sf %>% 
  sf::st_transform(sf::st_crs(lu_16)) %>% 
  sf::st_buffer(dist = 1000)

cams_sf_buffers_17 <- cams_sf %>% 
  sf::st_transform(sf::st_crs(lu_17)) %>% 
  sf::st_buffer(dist = 1000)

@@

events_plot_tb_filt_with_col <- events_plot_tb_filt %>% 
  tibble::rownames_to_column("plot_id")
```

>>>>

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '~/Documents/WILDLab/rofcamtrap/')
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
backup_options <- options()
options(scipen = 1, digits = 2)
```

@@

RoF_CameraDeploymentLog_WLU <- 
  readxl::read_excel(paste0("data-raw/tabular/RoF_CameraDeploymentLog.xlsx")) |> 
  janitor::clean_names()

cams_sf <- RoF_CameraDeploymentLog_WLU |> 
  dplyr::rename(longitude = x, latitude = y) |> 
  dplyr::filter(!is.na(latitude)) |> 
  dplyr::filter(!is.na(longitude)) |> 
  sf::st_as_sf(coords = c("longitude","latitude"), crs = "EPSG:4326", remove = F) |> 
  dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))

coords <- cams_sf |> 
  sf::st_coordinates() |> 
  as.data.frame()

@@

# zones <- 
#   sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
#               quiet=TRUE, ) |> 
#   sf::st_transform(crs = sf::st_crs(cams_sf))

zones <- 
  sf::st_read("data-raw/spatial/ecozones/EcoZone/EcoZone.shp",
              quiet = TRUE, ) |> 
  sf::st_transform(crs = sf::st_crs(cams_sf))

zones_filter <- zones |>
  sf::st_filter(cams_sf)

pols <- sf::st_coordinates(zones_filter) |> 
  as.data.frame()

# regions_labels <- zones_filter |> 
#   sf::st_drop_geometry() |> 
# dplyr::bind_cols(data.frame(x = c(-88.5, -84, -88.5),
#                             y = c(55, 52, 52.6)))

regions_labels <- zones_filter |> 
  sf::st_drop_geometry() |> 
  dplyr::bind_cols(data.frame(x = c(-84.5, -88.5),
                              y = c(52, 52.6)))

map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
                            zoom = 9, maptype = "stamen_terrain")
```

@@

base <- ggmap::ggmap(map) +
  ggplot2::labs(x = "Longitude", y = "Latitude") +
  # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
  ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2)

p <- base +
  # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
  #                     size = 2.5, data = regions_labels) +
  ggplot2::geom_label(ggplot2::aes(x = x, y = y, label = ECOZONE_NA), inherit.aes = FALSE,
                      size = 2.5, data = regions_labels) +
  ggplot2::geom_point(data = cams_sf_coords, #cex = 1, 
                      shape = 24, col = "black", fill = "red",
                      ggplot2::aes(x = X, y = Y, #fill=is_retrieved
                                   ), 
                      inherit.aes = FALSE) # +
  # ggplot2::labs(fill = "Camera \nRetrieved")
p
ggplot2::ggsave(plot = p, filename = "figures/camera_array.png", 
                width = 6, height = 8, dpi = 300)
ggplot2::ggsave(plot = p, filename = "reports/presentations/rofcamtrap-nov2023/images/camera_array.png", 
                width = 6, height = 8, dpi = 300)
```

@@

# base <- ggmap::ggmap(map) +
#   ggplot2::labs(x = "Longitude", y = "Latitude") +
#   # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
#   ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
# 
# p <- base +
#   # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
#   #                     size = 2.5, data = regions_labels) +
#   ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
#                       size = 2.5, data = regions_labels) +
#   ggplot2::geom_point(data = cams_sf_coords |> 
#                         dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
#                       col="black", fill = "red",
#                       ggplot2::aes(x=X, y=Y), 
#                       inherit.aes = FALSE) +
#   ggplot2::labs(fill = "Camera \nRetrieved")
# p
```

@@

anns_wide <- readRDS("data/objects/annotations_wide_treated.rds")
dets <- readRDS("data/objects/detections.rds")

@@

# Hardcoded check
stopifnot(nb_image == 3288178) # 2788375

exif <- readr::read_csv("data/tabular/all_exif.csv") |>
  janitor::clean_names() |>
  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) |>
  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |>

    dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |>

  dplyr::mutate(source_file = stringr::str_remove(source_file, 
                                               "/workspace/storage/TrailCamStorage_2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, 
                                               "/workspace/storage/TrailCamStorage/")) |>

  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |>

  dplyr::mutate(image_id = stringr::str_remove(image_id, 
                                               "_workspace_storage_TrailCamStorage_2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, 
                                               "_workspace_storage_TrailCamStorage_")) |>

  dplyr::mutate(image_id = strsplit(image_id, "_") |> 
                  purrr::map(~paste0(.x[2:length(.x)], collapse = "_")) |> 
                  unlist()) |>

  dplyr::mutate(
    source_file = 
      stringr::str_remove(source_file, 
                          stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
  dplyr::mutate(
    source_file = 
      stringr::str_remove(source_file, 
                          stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))
  # dplyr::mutate(image_id = stringr::str_sub(image_id, category6))
```

@@

# Select and distinct down to image-based infos
dets_images <- dets |>  dplyr::select(image_id, max_confidence, isempty) |> 
  dplyr::distinct()

# Breakdown of which images triggered a MD detection
dets_tb <- dets_images$isempty |> table()
dets_tb_sum <- sum(dets_tb)
dets_tb_percent <- dets_tb/dets_tb_sum*100

@@

```{r}
#| label: basics-stats-2
#| echo: false
dets_0 <- dets |> dplyr::filter(confidence > 0)
dets_0.1 <- dets |> dplyr::filter(confidence > 0.1)

# Breakdown of detections by categories (animal, person, vehicles)
dets_tb_cat_unfil <- dets_0 |> dplyr::pull(category_id) |> table()
dets_tb_cat_unfil_sum <- sum(dets_tb_cat_unfil)
dets_tb_cat_unfil_percent <- dets_tb_cat_unfil/dets_tb_cat_unfil_sum*100

@@

# Breakdown of detections above 0.1 by categories (animal, persons, vehicles)
dets_tb_cat <- dets_0.1 |> dplyr::pull(category_id) |> table()
dets_tb_cat_sum <- sum(dets_tb_cat)
dets_tb_cat_percent <- dets_tb_cat/sum(dets_tb_cat)*100

@@

# Histogram of confidence scores above 0.1
dets_0.1 |> 
  dplyr::mutate(category_id = as.character(category_id)) |> 
  dplyr::mutate(Type = dplyr::case_match(category_id,
                                         "0" ~ "Empty",
                                         "1" ~ "Animal",
                                         "2" ~ "Person",
                                         "3" ~ "Vehicle")) |> 
  ggplot2::ggplot() +
  ggplot2::theme_bw() +
  ggplot2::geom_histogram(ggplot2::aes(x = confidence, group = Type, fill = Type),
                          bins = 100) +
  ggplot2::scale_fill_viridis_d() +
  ggplot2::xlab("Confidence Score") +
  ggplot2::ylab("Count") +
  ggplot2::geom_vline(ggplot2::aes(xintercept = 0.1), lty = 2, col = "firebrick")
```

@@

# Number of uploaded images
uploaded <- dets |> dplyr::filter(confidence > 0.1) |> dplyr::pull(image_id) |> 
  unique() |> length()

@@

# Filter for non empty images
anns_wide_noempty <- anns_wide |>
  dplyr::filter(!is.na(to_name))

saveRDS(anns_wide_noempty, "data/objects/annotations_wide_noempty.rds")

@@

# Deleted predictions
dels <- anns_wide_noempty |> dplyr::filter(is.na(species))
dels_nb <- nrow(dels)
dels_nb_fmt <- prettyNum(dels_nb, big.mark = ",")
dels_nb_pct <- dels_nb/nrow(anns_wide_noempty)*100

# Should be no manuals here
stopifnot((anns_wide_noempty |> dplyr::filter(is.na(species)) 
           |> dplyr::filter(manual)) == 0)

# Actual annotations
true_anns <- anns_wide_noempty |> dplyr::filter(!is.na(species))
true_anns_nb <- nrow(true_anns)
# true_anns_nb_fmt <- prettyNum(true_anns_nb, big.mark = ",")
# true_anns_nb_pct <- true_anns_nb/processed*100

# Non manuals 
non_mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
  dplyr::filter(!manual)
non_mans_nb <- nrow(non_mans)
non_mans_nb_fmt <- prettyNum(non_mans_nb, big.mark = ",")
non_mans_nb_pct <- non_mans_nb/nrow(anns_wide_noempty)*100

# Mans
mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
  dplyr::filter(manual)
mans_nb <- nrow(mans)
mans_nb_fmt <- prettyNum(mans_nb, big.mark = ",")
mans_nb_pct <- mans_nb/nrow(anns_wide_noempty)*100

# adjusted
orig_tb <- anns_wide_noempty$origin |> table()
orig_tb_pct <- orig_tb/sum(orig_tb)*100
orig_tb_fmt <- prettyNum(orig_tb, big.mark = ",")
```

@@

# Only annotations for animals
# TODO !!!! investigate weird id of cut length 1
anns_wide_animals <- anns_wide_noempty |>
  dplyr::filter(species != 'Human') |> 
  dplyr::filter(species != 'Vehicle') |> 
  dplyr::mutate(id_cut_length = purrr::map_dbl(id , \(x) {
    x |> stringr::str_split("_") |> purrr::map_dbl(~length(.x))
  })) |> 
  dplyr::filter(id_cut_length > 1) |> 
  dplyr::mutate(camera_id = purrr::map_chr(id , \(x) {
    x |>
      stringr::str_split("_") |>
      purrr::map(~.x[[2]]) |>
      unlist()
  }))

# Unknowns and bird breakdown
# anns_wide_animals |> dplyr::filter(species == "Unknown") |> 
#   dplyr::pull(id) |> 
#   stringr::str_sub(1,4) |> table() |> sort() 
# anns_wide_animals |> dplyr::filter(stringr::str_detect(species,"Bird")) |> 
#   dplyr::pull(id) |> 
#   stringr::str_sub(1,4) |> table() |> sort()

@@

anns_wide_animals_joined <- anns_wide_animals |>
  dplyr::left_join(exif, by = c("source_file", "image_id"))

anns_wide_animals_joined_mut <- anns_wide_animals_joined |> 
  dplyr::filter(species != "Unknown") |>
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) |> 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_"))

# anns_wide_animals_joined_dets <- anns_wide_animals |> 
#   dplyr::right_join(dets_images, by = "image_id")
```

<!-- Other camera plots -->

```{r eval=FALSE}
# library(forcats)
library(ggplot2)
# anns_wide_animals_joined_mut |> 
#   ggplot(aes(x=date_time, y=species)) + geom_point()
# anns_wide_animals_joined_mut |> 
#   dplyr::filter(species == "Fisher") |> View()

camera_days <- exif |> dplyr::group_by(maker_notes_user_label) |> 
  dplyr::summarise(camera_first_time=min(exif_date_time_original)) |> 
  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_")) |> 
  dplyr::select(camera_id = maker_notes_user_label, 
                camera_first_day = date_time)

camera_regions <- cams_sf |> sf::st_join(zones) |> 
  sf::st_drop_geometry() |> dplyr::select(camera_id, ECOZONE_NA)

bp <- anns_wide_animals_joined_mut |>

  dplyr::filter(species != "Fisher") |> 
  dplyr::left_join(camera_days, by = "camera_id") |>

  dplyr::select(species, camera_id, 
                date_time, camera_first_day, ymd) |>

  dplyr::group_by(species, camera_id) |> 
  dplyr::summarise(first_day = min(date_time),
                   camera_first_day = min(camera_first_day),
                   camera_events = length(unique(ymd))) |> 
  dplyr::ungroup() |>

  dplyr::left_join(camera_regions, by = "camera_id") |>

  dplyr::group_by(species) |> 
  dplyr::mutate(total_events = sum(camera_events),
                n_cameras = dplyr::n()) |> 
  dplyr::ungroup() |>

  # dplyr::filter(n_cameras > 4) |>
  dplyr::filter(n_cameras > 3) |>

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) |>

@@

  scale_x_continuous()+
  labs(x = "Detection time (in days)",
       y = "Species",
       color = "Ecozone") + 
  theme_bw()

ggsave(plot = bp, filename = "figures/bp.png", width = 10, height = 15)

pts <- anns_wide_animals_joined_mut |>

  dplyr::filter(species != "Fisher") |> 
  dplyr::left_join(camera_days, by = "camera_id") |>

  dplyr::select(species, camera_id, 
                date_time, camera_first_day, ymd) |>

  dplyr::group_by(species, camera_id) |> 
  dplyr::summarise(first_day = min(date_time),
                   camera_first_day = min(camera_first_day),
                   camera_events = length(unique(ymd))) |> 
  dplyr::ungroup() |>

  dplyr::left_join(camera_regions, by = "camera_id") |>

  dplyr::group_by(species) |> 
  dplyr::mutate(total_events = sum(camera_events),
                n_cameras = dplyr::n()) |> 
  dplyr::ungroup() |>

  # dplyr::filter(n_cameras > 4) |>
  dplyr::filter(n_cameras > 3) |>

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) |>

@@

  scale_x_continuous()+
  labs(x = "Detection time (in days)",
       y = "Species",
       color = "Ecozone",
       size = "Day counts") + 
  theme_bw()
ggsave(plot = pts, filename = "figures/pts.png", width = 10, height = 15)
pts
```

```{r eval=FALSE}
# for_vic <- anns_wide_animals_joined_mut |> 
#   dplyr::select(id,species,date_time, camera_id) |> 
#   dplyr::left_join(cams_sf |> dplyr::select(camera_id, plot_id) |> sf::st_drop_geometry(),
#                    by = "camera_id")
# tot_camera <- for_vic$camera_id |> unique() |> length()
# hit_rates <- for_vic |> dplyr::select(species, camera_id) |> 
#   dplyr::group_by(species) |> dplyr::summarise(n_cameras=length(unique(camera_id))) |> 
#   dplyr::ungroup() |> dplyr::mutate(rate = n_cameras/tot_camera*100) |> 
#   dplyr::arrange(dplyr::desc(rate))
# # det_history_two_weeks <- 
# 
# readr::write_csv(for_vic, "../data/derived_data/observations_by_camera_2023Aug15.csv")
# readr::write_csv(hit_rates, "../data/derived_data/hit_rates_by_species_2023Aug15.csv")
```

```{r}
#| label: species
#| echo: false
anns_wide_animals_table <- anns_wide_animals |>
  dplyr::select(camera_id, species) |> 
  dplyr::filter(species != "Unknown") |>
  table() 
anns_wide_animals_table <- anns_wide_animals_table > 0
anns_wide_animals_table <- anns_wide_animals_table + 0
anns_wide_animals_table <- anns_wide_animals_table |> 
  as.data.frame() |> 
  tibble::rownames_to_column("camera_id") |> 
  dplyr::arrange(camera_id)
anns_wide_animals_table_tb <- anns_wide_animals_table |> 
  dplyr::select(-camera_id)
rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id
```

@@

lu_16 <- raster::raster("data-raw/spatial/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
lu_17 <- raster::raster("data-raw/spatial/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")

lu_dat <- readr::read_csv("data-raw/spatial/land_use/attr_table_northen_ont_lc.txt") |> 
  dplyr::mutate(cats = as.factor(code))

cams_sf_buffers_16 <- cams_sf |> 
  sf::st_transform(sf::st_crs(lu_16)) |> 
  sf::st_buffer(dist = 1000)

cams_sf_buffers_17 <- cams_sf |> 
  sf::st_transform(sf::st_crs(lu_17)) |> 
  sf::st_buffer(dist = 1000)

@@

events_plot_tb_filt_with_col <- events_plot_tb_filt |> 
  tibble::rownames_to_column("plot_id")
```
